---
title: "National patterns of alpha-gal sensitisation in Australia"
subtitle: "Laboratory surveillance of alpha-gal-specific IgE testing, 2014-2024"
format:
  html:
    page-layout: full
    theme:
      light: cosmo
      dark: darkly
    max-width: 1800px
    toc: true
    toc-depth: 2
    toc-location: left
    self-contained: true
    code-fold: true
    fig-width: 10
    fig-height: 6
---

```{r setup, include=FALSE}
# Load required packages
library(leaflet)
library(sf)
library(dplyr)
library(readr)
library(DT)
library(ggplot2)
library(tidyr)
library(plotly)
library(geojsonsf)
library(smoothr)
library(scales)
library(htmltools)
library(htmlwidgets)

# Suppress messages
options(warn = -1)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-data, cache=TRUE}
# ============================================================================
# DATA LOADING
# ============================================================================

# Load SA3 shapefile for map
sa3_sf <- st_read("public_data/sa3_cases_ESCat.shp", quiet = TRUE)

# Load tick distribution line
holocyclus_cache <- "public_data/holocyclus_line_processed.rds"
if (file.exists(holocyclus_cache)) {
  holocyclus_line <- readRDS(holocyclus_cache)
} else {
  holocyclus_line <- geojson_sf(
    "public_data/holocyclus_distribution_polyline.geojson"
  ) %>%
    filter(st_geometry_type(.) == "LINESTRING") %>%
    smooth(method = "spline")
  saveRDS(holocyclus_line, holocyclus_cache)
}

# Load CSV data
sa3_cases_csv <- read_csv("public_data/sa3_cases_ESCat.csv", show_col_types = FALSE)

# Figure 1 data: annual testing volume and its region-level decomposition
fig1_annual_tests <- read_csv("public_data/rev_fig1_annual_tests.csv", show_col_types = FALSE)
fig1_sa3_expansion <- read_csv("public_data/rev_fig1_sa3_expansion.csv", show_col_types = FALSE)
fig1_region_decomp <- read_csv("public_data/rev_fig1_region_decomp.csv", show_col_types = FALSE)
fig1_intensity_change <- read_csv("public_data/rev_fig1_intensity_change.csv", show_col_types = FALSE)

# Figure 2 data: detections, positivity, and case-growth decomposition
fig2_annual <- read_csv("public_data/rev_fig2_annual.csv", show_col_types = FALSE)
fig2_band_cases <- read_csv("public_data/rev_fig2_band_cases.csv", show_col_types = FALSE)
fig2_band_rate <- read_csv("public_data/rev_fig2_band_rate.csv", show_col_types = FALSE)
fig2_breakpoint <- read_csv("public_data/rev_fig2_breakpoint.csv", show_col_types = FALSE)
fig2_growth_decomp <- read_csv("public_data/rev_fig2_growth_decomp.csv", show_col_types = FALSE)

# Figure 4 data: geographic concentration
fig4_concentration <- read_csv("public_data/rev_fig4_concentration.csv", show_col_types = FALSE)

# Figure 5 data: longitudinal alpha-Gal sIgE
fig5_slopes <- read_csv("public_data/rev_fig5_slopes.csv", show_col_types = FALSE)
fig5_first_last <- read_csv("public_data/rev_fig5_first_last.csv", show_col_types = FALSE)

# Supplementary figure data
figS1_hist <- read_csv("public_data/figS1_agal_hist.csv", show_col_types = FALSE)
figS2_all <- read_csv("public_data/figS2_age_sex_all.csv", show_col_types = FALSE)
figS2_pos <- read_csv("public_data/figS2_age_sex_pos.csv", show_col_types = FALSE)
rr_sex_by_age <- read_csv("public_data/rr_sex_by_age.csv", show_col_types = FALSE)
rr_age_only <- read_csv("public_data/rr_age_only.csv", show_col_types = FALSE)
combined_rr <- read_csv("public_data/combined_rr_results.csv", show_col_types = FALSE)
figS4_rules <- read_csv("public_data/figS4_attribution_rules.csv", show_col_types = FALSE)
figS5_distances <- read_csv("public_data/figS5_distances.csv", show_col_types = FALSE)
figS6_supp <- read_csv("public_data/figS6_supp.csv", show_col_types = FALSE)
figS6_lorenz <- read_csv("public_data/figS6_lorenz.csv", show_col_types = FALSE)
figS6_gini <- read_csv("public_data/figS6_gini.csv", show_col_types = FALSE)
```

```{r colors}
# ============================================================================
# COLOR SCHEME
# ============================================================================

MidnightBlue <- "#00313C"
MiddayBlue <- "#00A9CE"
STEEL <- "#757579"
MIST <- "#DADBDC"
GOLD <- "#FFB81C"
PLUM <- "#6D2077"
BLUEBERRY <- "#1F22AA"
TEAL <- "#017377"
ORANGE <- "#E87721"
MALE_COL <- "#00A9CE"
FEMALE_COL <- "#E4002B"

# Trend-panel colours, matching the published figures
LINE_COL <- "#18444E"
POINT_COL <- "#48575F"
TREND_COL <- "red"

# Map color mapping
color_mapping_blue <- c(
  ">300" = "#08306b",
  "100-299" = "#08519c",
  "30-99" = "#4292c6",
  "10-29" = "#9ecae1",
  "<10" = "#deebf7",
  "0" = "#f7f7f7",
  "No data" = "grey90"
)

category_order <- c(
  ">300",
  "100-299",
  "30-99",
  "10-29",
  "<10",
  "0",
  "No data"
)

# Age categories order
age_order <- c("0-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75-84", "85+")
```

```{r prepare-spatial}
# ============================================================================
# PREPARE SPATIAL DATA
# ============================================================================

# Transform to WGS84 for Leaflet
sa3_sf <- st_transform(sa3_sf, crs = 4326)

# Ensure ESCat is a factor with correct order.
# The shapefile stores the bands with a trailing " Suspected cases" (e.g.
# "<10 Suspected cases"), which the labels above no longer use. Strip the
# suffix before factoring: matching on the raw strings would send every region
# to NA and render the whole map in the no-data colour.
escat_raw <- as.character(sa3_sf$ESCat)
sa3_sf$ESCat <- factor(
  sub("[[:space:]]+Suspected cases$", "", escat_raw),
  levels = category_order
)

# Fail loudly rather than quietly drawing a grey map if the two ever diverge
unmatched <- unique(escat_raw[!is.na(escat_raw) & is.na(sa3_sf$ESCat)])
if (length(unmatched) > 0) {
  stop(
    "ESCat values in the shapefile are not in category_order: ",
    paste(unmatched, collapse = ", ")
  )
}

# Create color palette function
pal <- colorFactor(
  palette = color_mapping_blue,
  domain = category_order,
  ordered = TRUE
)
```

```{r captions}
# ============================================================================
# FIGURE CAPTIONS
# Held in one place so each panel of a multi-panel figure repeats the same text.
# ============================================================================

fig1_caption <- paste0(
  "**Figure 1** Annual trends in α-Gal sIgE testing and its decomposition, ",
  "2014-2024. **(A)** Total tests performed. **(B)** Number of SA3 regions ",
  "from which α-Gal sIgE test was performed. **(C)** Sources of the increase ",
  "in testing volume relative to 2014, decomposed across SA3 regions into ",
  "tests performed in regions new since 2014, the net change within regions ",
  "testing in both years, and tests lost from regions that ceased testing. ",
  "Circles mark the observed net change. **(D)** The percentage change in ",
  "testing intensity within SA3 regions; points show individual SA3 regions, ",
  "violin plot shows probability density, boxplot indicates median (63.0%) ",
  "and interquartile range. Red dashed lines and ribbons indicate linear ",
  "regression and standard errors, respectively. sIgE: specific ",
  "immunoglobulin E."
)

fig2_caption <- paste0(
  "**Figure 2** Annual α-Gal sensitisation detections and test positivity, ",
  "2014-2024. **(A)** Number of α-Gal sensitised individuals detected ",
  "annually with segmented linear regression model fit. **(B)** Annual ",
  "positivity rate with linear regression model fit, excluding outlier year ",
  "2014. **(C)** Decomposition of the growth in detected α-Gal sensitisation ",
  "relative to the pooled 2018-2020 reference period. Segmented and linear ",
  "regression models and standard errors indicated by red dashed lines and ",
  "ribbons, respectively. Dotted line and light blue shading indicates the ",
  "segmented linear regression breakpoint with 95% confidence intervals. ",
  "sIgE: specific immunoglobulin E."
)

fig3_caption <- paste0(
  "**Figure 3** Geographic distribution of laboratory-detected α-Gal ",
  "sIgE-positive individuals (by residential SA3) per 1M PPY in Australia, ",
  "2014-2024"
)

fig4_caption <- paste0(
  "**Figure 4** Geographic concentration of α‑Gal sensitisation across ",
  "Australian SA3 regions. **(A)** Lorenz curve showing cumulative ",
  "distribution of sensitisation across SA3 regions (ranked by case burden). ",
  "The diagonal dashed line represents perfect equality; deviation from this ",
  "line indicates concentration. **(B)** Pareto chart showing the proportion ",
  "of total sensitisation accounted for by cumulative percentages of SA3 ",
  "regions. PPY: population per year."
)

fig5_caption <- paste0(
  "**Figure 5** Longitudinal changes in α-Gal sIgE levels in individuals ",
  "with ≥2 tests (n=1,017). **(A)** Distribution of individual annual rates ",
  "of change (β) in log10 α-Gal sIgE/year percentage change from previous ",
  "test, estimated from linear mixed-effects models. Dashed line indicates ",
  "the median rate of change. **(B)** Comparison of first and last α-Gal ",
  "sIgE measurements for each individual. The red dashed lines and shaded ",
  "area show the linear regression and standard error, respectively. α-Gal ",
  "sIgE levels reported at actual values below the 0.1 kU/L positivity ",
  "cut-off were retained as real values and those reported as ",
  "“\\<0.1 kU/L” were assigned 0.1 kU/L but flagged as negative test ",
  "results; 17 undetectable results were retained at zero and 13 results ",
  "reported as “\\>100 kU/L” were assigned 100 kU/L. sIgE: specific ",
  "immunoglobulin E."
)

figS1_caption <- paste0(
  "**Figure S1** Histogram of the distribution of α‑Gal sIgE tests results ",
  "across the whole tested cohort. Red dotted line shows the 0.1 kU/L ",
  "positivity threshold. sIgE: specific immunoglobulin E."
)

figS2_caption <- paste0(
  "**Figure S2** Age distribution by sex of all tested people **(A)** and ",
  "people positive (0.1 kU/L) α‑Gal sIgE tests."
)

figS3_caption <- paste0(
  "**Figure S3** Positivity ratios comparing: **(A)** males to females within ",
  "each age category, **(B)** each age group to the reference (25–34 years, ",
  "pooled sexes), and **(C)** all age-sex combinations to females aged 25–34 ",
  "years (the lowest-positivity group, 18.5% positivity). Positivity ratios ",
  "represent the ratio of test positivity between strata among tested ",
  "individuals and do not indicate population-level risk. Filled circles ",
  "indicate statistical significance (p \\< 0.01); open circles indicate ",
  "non-significant comparisons. Error bars represent 95% confidence ",
  "intervals. Dashed vertical line indicates a positivity ratio of 1.0 (no ",
  "difference from reference group)."
)

figS4_caption <- paste0(
  "**Figure S4** Share of observed 2014–2024 growth in α-gal sIgE testing ",
  "volume attributed to geographic expansion and to testing intensity, under ",
  "six decompositions of the identity T=N×I. The attributed geographic share ",
  "ranges from 26% to 56% depending only on how the cross-product is handled."
)

figS5_caption <- paste0(
  "**Figure S5** Geographic distance between patient residential and ",
  "referring clinicians practice SA2 region showing how far patients travel ",
  "for their consultation for 6,404 tests in which both variables were ",
  "obtained."
)

figS6_caption <- paste0(
  "**Figure S6** Geographic concentration of α-Gal sensitisation standardised ",
  "by testing volume, Australian SA3 regions, 2014–2024. **(A)** Lorenz ",
  "curves for sensitised individuals and persons tested, with population as ",
  "denominator, regions ranked from highest to lowest rate; the dashed ",
  "diagonal is the line of equality. The curve for sensitisation lies outside ",
  "that for testing, indicating greater concentration. **(B)** Lorenz curve ",
  "for sensitised individuals with persons tested as denominator, regions ",
  "ranked by positivity, removing population from the comparison (Gini ",
  "coefficient = 0.30). **(C)** Standardised positivity ratios with Poisson ",
  "95% confidence intervals for the 30 regions with the highest ",
  "population-adjusted burden, ordered by ratio; the dashed line marks parity ",
  "with national positivity and point size is proportional to persons tested. ",
  "**(D)** Sensitised individuals per 1M PPY against persons tested per 1M ",
  "PPY, logarithmic axes, point size proportional to persons tested; the ",
  "dashed line is the expectation under constant national positivity (35.6%), ",
  "and the 30 regions with the highest population-adjusted burden are ",
  "highlighted. PPY: population per year."
)

figS7_caption <- paste0(
  "**Figure S7** Population-standardised and testing-standardised burden of ",
  "α-Gal sensitisation, Australian SA3 regions, 2014–2024. Each panel shows ",
  "all of mainland Australia (left) and zoomed in south-east Queensland and ",
  "eastern New South Wales (right). **(A)** Sensitised individuals per ",
  "million person-years, the population-standardised rate reported in the ",
  "main text. **(B)** Persons tested per million person-years, showing ",
  "testing intensity. **(C)** Test positivity (empirical Bayes smoothed) for ",
  "regions with at least 20 persons tested, in which testing volume is the ",
  "denominator; national positivity is 35.6% and regions with insufficient ",
  "testing are shown in grey. Colour scales in A and B are square-root ",
  "transformed."
)
```

# Disclaimer {.unnumbered}

**Preprint Research Notice:** The findings and analyses displayed herein are part of ongoing research and are provisional. They may be *incomplete, subject to revision, or contain errors*.

# Home {.unnumbered}



## Published Manuscript

[Link to published manuscript](https://doi.org/10.XXXX/XXXXXX)

## Abstract

**Objectives:** To characterise national patterns of alpha-gal (α‑Gal) specific immunoglobulin E (sIgE) sensitisation in Australia, including temporal trends in testing and detection, geographic distribution, demographic correlates of test positivity, and longitudinal antibody dynamics.

**Design:** Retrospective analysis of α‑Gal sIgE ImmunoCAP™ test results from 1 January 2014 to 31 December 2024.

**Setting, Participants:** Deidentified laboratory records from 14,075 individuals tested across Australia, with residential postcodes mapped to Statistical Areas Level 3 (SA3).

**Main outcome measures:** Test volumes, α-Gal sensitisation (positive test) counts, positivity rates, spatial clustering metrics, demographic correlates of test positivity, and longitudinal α‑Gal sIgE trajectories.

**Results:** Overall, 35.7% (5,025) of individuals tested positive (≥0.1 kU/L α‑Gal sIgE). Testing volume increased 330% over the study period, with detections accelerating from around 2022. Decomposition attributed 79% of case growth to expanded testing, with up to 21%unexplained by measurable surveillance changes. Sensitised individuals clustered along the eastern seaboard within the range of Ixodes holocyclus, with extreme spatial clustering: three SA3 regions accounted for over one-quarter of nationally detected sensitisation. Females comprised the majority of sensitised individuals, although males had higher positivity. Positivity increased with age, peaking at 45-74 years. Among 1,017 repeatedly tested individuals, α‑Gal sIgE levels antibodies declined in 73.7% of people by a median of 20.4% annually.

**Conclusions:** This first national assessment of α‑Gal sIgE sensitisation in Australia reveals a substantial, geographically concentrated, and growing burden. The spatial concentration suggests targeted public health interventions could efficiently reach most highest-risk people. Testing expansion is the dominant driver of rising case numbers, but a residual increase remains that surveillance expansion alone does not explain. Laboratory data cannot separate this from rising clinician and public awareness, and prospective studies are needed to confirm genuine disease emergence.

## About This Interactive Application

This interactive web application provides access to the data visualizations in the [National patterns of alpha-gal sensitisation in Australia: Laboratory surveillance of alpha-gal-specific IgE testing, 2014-2024]((https://doi.org/10.XXXX/XXXXXX)) (PUBLICATION DETAILS).

Users can explore:

-   Interactive versions of every manuscript figure, panel by panel, with zoom, pan and hover detail
-   **Figure S7** is the one exception, shown as published: each panel pairs a national map with an east-coast inset, and Figure 3 above is the zoomable equivalent
-   **Data tables and detailed regional statistics**

Navigate through the sections below to explore different aspects of the analysis. All figures are interactive and can be zoomed and panned for detailed examination.

## Data Sources

Deidentified data were obtained from all [a-Gal sIgE immunoCAP™](https://www.allergy.org.au/patients/food-allergy/mammalian-meat-tick-faq?highlight=WyJ0byJd) tests submitted to the following pathology service providers between January 1 2014, and December 31 2024:

-   [QML Pathology](https://www.qml.com.au/)
-   [Sullivan Nicolaides Pathology](https://www.snp.com.au/)
-   [Douglas Hanley Moir Pathology](https://www.dhm.com.au/)
-   [Laverty Pathology](https://www.laverty.com.au/)

Patients’ residential locations were aggregated to Australian Bureau of Statistics (ABS) statistical areas level 3 (SA3) define by the [Australian Statistical Geography Standard Edition 3](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/latest-release), which is designed to create meaningful functional areas of regional towns and cities with populations between 30,000 and 130,000.

Adjusted number of sensitised individuals are displayed as cases per 1M population per year (1M PPY) which adjusts for both regional population size and samples time from (2014-2024) using annual regional population estimates from the [ABS](https://www.abs.gov.au/statistics/people/population/national-state-and-territory-population/latest-release) as population denominators.

```
Sensitised individuals per 1M PPY = (cumulative_incidence_2014-2024 / cumulative_annual_population_2014-2024) * 1000000
```

------------------------------------------------------------------------

© 2025 \| For questions or feedback, please contact: alexander.gofton\@csiro.au


# Figure 1. Annual trends in α-Gal sIgE testing and its decomposition, 2014-2024

::: panel-tabset
## A. Total tests

```{r fig1a}
tests_model <- lm(total_tests ~ Year, data = fig1_annual_tests)
tests_pred <- data.frame(
  Year = fig1_annual_tests$Year,
  fit = predict(tests_model, se.fit = TRUE)$fit,
  se = predict(tests_model, se.fit = TRUE)$se.fit
) %>%
  mutate(lower = fit - 1.96 * se, upper = fit + 1.96 * se)

plot_1a <- ggplot(fig1_annual_tests, aes(x = Year, y = total_tests)) +
  geom_ribbon(
    data = tests_pred,
    aes(y = fit, ymin = lower, ymax = upper),
    fill = TREND_COL,
    alpha = 0.3
  ) +
  geom_line(
    data = tests_pred,
    aes(y = fit),
    colour = TREND_COL,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_line(colour = LINE_COL, linewidth = 0.6) +
  geom_point(colour = LINE_COL, fill = POINT_COL, size = 3, shape = 21) +
  scale_x_continuous(breaks = seq(2014, 2024, 1)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, NA),
    labels = comma
  ) +
  labs(x = NULL, y = "α-Gal sIgE tests") +
  theme_minimal()

ggplotly(plot_1a, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 40, t = 20))
```

`r fig1_caption`

## B. Testing regions

```{r fig1b}
geo_model <- lm(n_sa3_regions ~ YOT, data = fig1_sa3_expansion)
geo_pred <- data.frame(
  YOT = fig1_sa3_expansion$YOT,
  fit = predict(geo_model, se.fit = TRUE)$fit,
  se = predict(geo_model, se.fit = TRUE)$se.fit
) %>%
  mutate(lower = fit - 1.96 * se, upper = fit + 1.96 * se)

plot_1b <- ggplot(fig1_sa3_expansion, aes(x = YOT, y = n_sa3_regions)) +
  geom_ribbon(
    data = geo_pred,
    aes(y = fit, ymin = lower, ymax = upper),
    fill = TREND_COL,
    alpha = 0.3
  ) +
  geom_line(
    data = geo_pred,
    aes(y = fit),
    colour = TREND_COL,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_line(colour = LINE_COL, linewidth = 0.6) +
  geom_point(colour = LINE_COL, fill = POINT_COL, size = 3, shape = 21) +
  scale_x_continuous(breaks = seq(2014, 2024, 1)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, NA)
  ) +
  labs(x = NULL, y = "Number of SA3 regions") +
  theme_minimal()

ggplotly(plot_1b, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 40, t = 20))
```

`r fig1_caption`

## C. Sources of testing growth

```{r fig1c}
decomp_levels <- c(
  "Regions new since 2014",
  "Change within continuing regions",
  "Regions that ceased testing"
)

decomp_long <- fig1_region_decomp %>%
  select(
    Year,
    `Regions new since 2014` = tests_entrant,
    `Regions that ceased testing` = tests_exiting,
    `Change within continuing regions` = tests_continuing
  ) %>%
  pivot_longer(-Year, names_to = "component", values_to = "tests") %>%
  mutate(component = factor(component, levels = decomp_levels))

plot_1c <- ggplot(decomp_long, aes(x = Year, y = tests, fill = component)) +
  geom_col(width = 0.9, alpha = 0.7) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey30") +
  geom_point(
    data = fig1_region_decomp,
    aes(x = Year, y = dT),
    inherit.aes = FALSE,
    shape = 21,
    fill = MIST,
    colour = "black",
    size = 2.5
  ) +
  scale_fill_manual(
    values = c(
      "Regions new since 2014" = MiddayBlue,
      "Change within continuing regions" = MidnightBlue,
      "Regions that ceased testing" = STEEL
    )
  ) +
  scale_x_continuous(breaks = seq(2015, 2024, 1)) +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Additional tests relative to 2014", fill = NULL) +
  theme_minimal()

ggplotly(plot_1c, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.18),
    margin = list(l = 80, r = 40, b = 90, t = 20)
  )
```

`r fig1_caption`

## D. Within-region testing intensity

```{r fig1d}
intensity_median <- median(fig1_intensity_change$intensity_pct_change)

plot_1d <- ggplot(
  fig1_intensity_change,
  aes(x = "", y = intensity_pct_change)
) +
  geom_violin(fill = MiddayBlue, alpha = 0.3, linewidth = 0.3) +
  geom_jitter(
    aes(text = paste0("SA3 ", SA3_code, ": ", round(intensity_pct_change, 1), "%")),
    width = 0.25,
    alpha = 0.7,
    size = 1,
    colour = MidnightBlue,
    shape = 16
  ) +
  geom_boxplot(
    width = 0.15,
    fill = "white",
    alpha = 1,
    linewidth = 0.3,
    colour = "black",
    outlier.shape = NA
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = "SA3 regions", y = "Testing intensity change (%)") +
  theme_minimal() +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggplotly(plot_1d, tooltip = "text") %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 40, t = 20))
```

`r fig1_caption`
:::

# Figure 2. Annual α-Gal sensitisation detections and test positivity, 2014-2024

::: panel-tabset
## A. Sensitised individuals detected

```{r fig2a}
plot_2a <- ggplot(fig2_annual, aes(x = Year)) +
  annotate(
    "rect",
    xmin = fig2_breakpoint$lower,
    xmax = fig2_breakpoint$upper,
    ymin = 0,
    ymax = max(fig2_band_cases$upper, fig2_annual$positive_tests) * 1.08,
    fill = MiddayBlue,
    alpha = 0.12
  ) +
  geom_vline(
    xintercept = fig2_breakpoint$psi,
    linetype = "dotted",
    colour = "black",
    alpha = 0.8,
    linewidth = 0.4
  ) +
  geom_ribbon(
    data = fig2_band_cases,
    aes(y = fit, ymin = lower, ymax = upper),
    fill = TREND_COL,
    alpha = 0.3
  ) +
  geom_line(
    data = fig2_band_cases,
    aes(y = fit),
    colour = TREND_COL,
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  geom_line(aes(y = positive_tests), colour = LINE_COL, linewidth = 0.6) +
  geom_point(
    aes(y = positive_tests),
    colour = LINE_COL,
    fill = POINT_COL,
    size = 3,
    shape = 21
  ) +
  scale_x_continuous(breaks = seq(2014, 2024, 1)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.08)),
    limits = c(0, NA),
    labels = comma
  ) +
  labs(x = NULL, y = "α-Gal sensitised individuals detected") +
  theme_minimal()

ggplotly(plot_2a, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 40, t = 20))
```

`r fig2_caption`

## B. Positivity rate

```{r fig2b}
plot_2b <- ggplot(fig2_annual, aes(x = Year)) +
  geom_ribbon(
    data = fig2_band_rate,
    aes(y = fit, ymin = lower, ymax = upper),
    fill = TREND_COL,
    alpha = 0.3
  ) +
  geom_line(
    data = fig2_band_rate,
    aes(y = fit),
    colour = TREND_COL,
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  geom_line(aes(y = pos_prop), colour = LINE_COL, linewidth = 0.6) +
  geom_point(
    aes(y = pos_prop),
    colour = LINE_COL,
    fill = POINT_COL,
    size = 3,
    shape = 21
  ) +
  scale_x_continuous(breaks = seq(2014, 2024, 1)) +
  scale_y_continuous(
    limits = c(0, NA),
    labels = function(x) paste0(round(x, 1), "%")
  ) +
  labs(x = NULL, y = "Positivity rate (%)") +
  theme_minimal()

ggplotly(plot_2b, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 40, t = 20))
```

`r fig2_caption`

## C. Sources of detection growth

```{r fig2c}
growth_levels <- c(
  "Increased testing volume",
  "Changed composition of who was tested",
  "Not explained by either"
)

growth_long <- fig2_growth_decomp %>%
  select(
    Year,
    `Increased testing volume` = volume,
    `Changed composition of who was tested` = composition,
    `Not explained by either` = residual
  ) %>%
  pivot_longer(-Year, names_to = "component", values_to = "cases") %>%
  mutate(component = factor(component, levels = growth_levels)) %>%
  arrange(Year, component) %>%
  group_by(Year) %>%
  mutate(mid = cumsum(cases) - cases / 2, pct = cases / sum(cases) * 100) %>%
  ungroup()

plot_2c <- ggplot(growth_long, aes(x = Year, y = cases, fill = component)) +
  geom_col(
    width = 0.72,
    alpha = 0.7,
    position = position_stack(reverse = TRUE)
  ) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.3) +
  geom_text(
    data = growth_long %>% filter(cases >= 30),
    aes(y = mid, label = sprintf("%.0f%%", pct)),
    size = 3,
    colour = "white",
    fontface = "bold"
  ) +
  geom_text(
    data = fig2_growth_decomp,
    aes(x = Year, y = obs, label = sprintf("%+.0f", obs)),
    inherit.aes = FALSE,
    vjust = -0.6,
    size = 3.2,
    colour = MidnightBlue,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Increased testing volume" = MiddayBlue,
      "Changed composition of who was tested" = GOLD,
      "Not explained by either" = PLUM
    )
  ) +
  scale_x_continuous(breaks = fig2_growth_decomp$Year) +
  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0.03, 0.1))
  ) +
  labs(x = NULL, y = "Additional detections vs 2018-2020", fill = NULL) +
  theme_minimal()

ggplotly(plot_2c, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.18),
    margin = list(l = 80, r = 40, b = 90, t = 20)
  )
```

`r fig2_caption`
:::

# Figure 3. Geographic distribution of α-Gal sensitisation {#interactive-map}

```{r map, fig.height=8}
#| column: screen-inset

leaflet(sa3_sf, width = "100%", height = "800px") %>%
  # Add base map tiles. The CartoDB Positron and DarkMatter layers were dropped
  # because they now prompt for an API key; OpenStreetMap needs none, so there
  # is no longer a base layer to switch between.
  addProviderTiles(providers$OpenStreetMap, group = "OpenStreetMap") %>%
  # Create a custom pane for the holocyclus line with higher z-index
  addMapPane("holocyclus_pane", zIndex = 450) %>%
  # Add SA3 polygons
  addPolygons(
    fillColor = ~ pal(ESCat),
    fillOpacity = 0.7,
    color = "black",
    weight = 1,
    opacity = 1,
    group = "SA3 Regions",
    label = ~ paste0(
      SA_name,
      "<br>Sensitised individuals per 1M PPY: ",
      round(c1MPPY, 1),
      "<br>Total sensitised individuals: ",
      ct24pos
    ) %>% lapply(htmltools::HTML),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    ),
    highlightOptions = highlightOptions(
      weight = 3,
      color = "#666",
      fillOpacity = 0.9,
      bringToFront = TRUE
    )
  ) %>%
# Add Ixodes holocyclus distribution boundary line to custom pane
  addPolylines(
    data = holocyclus_line,
    color = "red",
    weight = 2.5,
    opacity = 0.9,
    dashArray = "10, 5",
    labelOptions = labelOptions(
      style = list("font-weight" = "bold"),
      textsize = "12px"
    ),
    label = "Ixodes holocyclus western distribution limit",
    group = "I. holocyclus distribution",
    options = pathOptions(pane = "holocyclus_pane")
  ) %>%

  addLegend(
    position = "bottomright",
    pal = pal,
    values = ~ ESCat,
    title = "Number of α-Gal sensitised<br>individuals per 1M PPY",
    opacity = 1
  ) %>%

  # Add custom legend entry for I. holocyclus distribution line
  addControl(
    html = paste0(
      '<div style="background: white; padding: 6px 10px; ',
      'font-size: 12px; line-height: 18px;">',
      '<span style="display: inline-block; width: 30px; height: 0; ',
      'border-top: 2.5px dashed red; vertical-align: middle; ',
      'margin-right: 8px;"></span>',
      'Approximate western edge<br>of I. holocyclus range<br>',
      '</div>'
    ),
    position = "bottomleft"
  ) %>%

  setView(lng = 133.7751, lat = -25.2744, zoom = 4) %>%

  # Add opacity slider control
  addControl(
    html = paste0(
      '<div style="background: white; padding: 8px 12px; ',
      'border-radius: 4px; box-shadow: 0 1px 5px rgba(0,0,0,0.4); ',
      'font-size: 12px; line-height: 20px;">',
      '<b>Polygon Opacity</b><br>',
      '<input id="opacitySlider" type="range" min="0" max="100" value="70" ',
      'style="width: 120px; vertical-align: middle;">',
      ' <span id="opacityValue">70%</span>',
      '</div>'
    ),
    position = "topright"
  ) %>%

  # Wire up the slider to control polygon fill opacity
  htmlwidgets::onRender("
    function(el, x) {
      var map = this;
      var slider = document.getElementById('opacitySlider');
      var label = document.getElementById('opacityValue');
      slider.addEventListener('input', function() {
        var opacity = this.value / 100;
        label.textContent = this.value + '%';
        map.eachLayer(function(layer) {
          if (layer.options && layer.options.group === 'SA3 Regions') {
            layer.setStyle({fillOpacity: opacity});
          }
        });
      });
    }
  ")
```

`r fig3_caption`

## Data Table

```{r table}
#| column: page

display_data <- sa3_cases_csv %>%
  select(SA_name, ct24pos, c1MPPY) %>%
  arrange(desc(c1MPPY))

datatable(
  display_data,
  colnames = c(
    "SA3 Region" = "SA_name",
    "α-Gal sensitised individuals (2014-2024)" = "ct24pos",
    "α-Gal sensitised individuals per 1M PPY" = "c1MPPY"
  ),
  options = list(
    pageLength = 15,
    scrollX = TRUE,
    order = list(list(2, "desc"))
  ),
  rownames = FALSE
) %>%
  formatRound("α-Gal sensitised individuals per 1M PPY", 1)
```

# Figure 4. Geographic concentration of α-Gal sensitisation

::: panel-tabset
## A. Lorenz curve

```{r fig4a}
library(patchwork)
library(forcats)

conc_theme <- theme_minimal() +
  theme(
    axis.text = element_text(size = 9),
    axis.title = element_text(size = 10)
  )

plot_4a <- fig4_concentration %>%
  ggplot(aes(x = regions_percentage, y = cumulative_percentage)) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "red",
    linewidth = 0.4
  ) +
  geom_line(colour = LINE_COL, linewidth = 0.5) +
  geom_point(
    aes(
      text = paste0(
        SA_name, " (", State, ")",
        "<br>", round(regions_percentage, 1), "% of SA3 regions",
        "<br>", round(cumulative_percentage, 1), "% of total sensitisation"
      )
    ),
    colour = LINE_COL,
    fill = POINT_COL,
    size = 1.8,
    shape = 21
  ) +
  scale_x_continuous(
    breaks = seq(0, 100, 20),
    limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    breaks = seq(0, 100, 25),
    limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  labs(
    x = "Percentage of SA3 regions",
    y = "Cumulative percentage of total α-Gal sensitisation per 1M PPY"
  ) +
  conc_theme

ggplotly(plot_4a, tooltip = "text") %>%
  layout(showlegend = FALSE, margin = list(l = 90, r = 40, b = 60, t = 20))
```

`r fig4_caption`

## B. Pareto chart

```{r fig4b}
# The published panel carries State as a coloured track beneath the axis, which
# has no interactive equivalent; here the same information colours the bars.
state_cols <- c(NSW = "#5FA2EB", QLD = "#73182C")

pareto_data <- fig4_concentration %>%
  head(30) %>%
  mutate(
    SA_name = fct_reorder(SA_name, -c1MPPY),
    State = factor(
      ifelse(is.na(State) | State == "", "Other", State),
      levels = c("NSW", "QLD", "VIC", "SA", "WA", "TAS", "NT", "ACT", "Other")
    )
  )

pareto_ymax <- max(pareto_data$c1MPPY)

plot_4b <- ggplot(pareto_data, aes(x = SA_name)) +
  geom_col(
    aes(
      y = c1MPPY,
      fill = State,
      text = paste0(
        SA_name, " (", State, ")",
        "<br>", round(c1MPPY, 1), " sensitised per 1M PPY",
        "<br>cumulative ", round(cumulative_percentage, 1), "%"
      )
    ),
    alpha = 0.8
  ) +
  geom_line(
    aes(y = cumulative_percentage * pareto_ymax / 100, group = 1),
    colour = LINE_COL,
    linewidth = 0.5
  ) +
  geom_point(
    aes(
      y = cumulative_percentage * pareto_ymax / 100,
      text = paste0(
        SA_name,
        "<br>cumulative ", round(cumulative_percentage, 1), "% of total"
      )
    ),
    colour = LINE_COL,
    fill = POINT_COL,
    size = 1.8,
    shape = 21
  ) +
  scale_fill_manual(values = state_cols, drop = TRUE, name = NULL) +
  scale_y_continuous(
    name = "α-Gal sensitisation per 1M PPY",
    limits = c(0, pareto_ymax),
    expand = c(0, 0)
  ) +
  labs(x = "SA3 regions ranked by number of cases per 1M PPY") +
  conc_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

# The right-hand cumulative axis is added at the plotly layer: ggplotly drops
# sec_axis(). An invisible trace anchors it, because plotly.js does not draw an
# overlaying axis that carries no data.
ggplotly(plot_4b, tooltip = "text") %>%
  add_trace(
    x = levels(pareto_data$SA_name)[1],
    y = 0,
    yaxis = "y2",
    type = "scatter",
    mode = "markers",
    marker = list(opacity = 0),
    showlegend = FALSE,
    hoverinfo = "skip",
    inherit = FALSE
  ) %>%
  layout(
    yaxis = list(range = c(0, pareto_ymax)),
    yaxis2 = list(
      overlaying = "y",
      side = "right",
      range = c(0, 100),
      ticksuffix = "%",
      showgrid = FALSE,
      zeroline = FALSE,
      title = list(
        text = "Cumulative percentage of total α-Gal sensitisation",
        font = list(size = 12)
      )
    ),
    legend = list(
      orientation = "v",
      x = 1.10,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 90, r = 160, b = 170, t = 20)
  )
```

`r fig4_caption`
:::

# Figure 5. Longitudinal changes in α-Gal sIgE levels

::: panel-tabset
## A. Annual rate of change

```{r fig5a}
median_pct <- median(fig5_slopes$annual_pct_change)

plot_5a <- ggplot(fig5_slopes, aes(x = annual_pct_change)) +
  geom_histogram(
    bins = 30,
    fill = "grey50",
    alpha = 0.7,
    colour = "grey10",
    linewidth = 0.3
  ) +
  geom_vline(
    xintercept = median_pct,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.6
  ) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Annual percentage change in α-Gal sIgE",
    y = "Number of people"
  ) +
  theme_minimal()

ggplotly(plot_5a, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 60, t = 20))
```

`r fig5_caption`

## B. First vs last test

```{r fig5b}
AGAL_OFFSET <- 0.1
AGAL_TICKS <- c(0, 0.1, 1, 10, 100)
agal_trans <- function(x) log10(x + AGAL_OFFSET)

first_last <- fig5_first_last %>%
  mutate(Change = ifelse(decreased, "Decreased", "Increased"))

plot_5b <- ggplot(first_last, aes(x = first_log_aGal, y = last_log_aGal)) +
  geom_point(aes(colour = Change), alpha = 0.6, size = 1.6, shape = 16) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    colour = TREND_COL,
    fill = TREND_COL,
    alpha = 0.3,
    linetype = "dashed",
    linewidth = 0.5
  ) +
  scale_colour_manual(
    values = c("Increased" = MidnightBlue, "Decreased" = MiddayBlue),
    name = NULL
  ) +
  scale_x_continuous(breaks = agal_trans(AGAL_TICKS), labels = AGAL_TICKS) +
  scale_y_continuous(breaks = agal_trans(AGAL_TICKS), labels = AGAL_TICKS) +
  labs(
    x = "First test α-Gal sIgE (kU/L)",
    y = "Last test α-Gal sIgE (kU/L)"
  ) +
  theme_minimal()

ggplotly(plot_5b, tooltip = c("x", "y", "colour")) %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 80, r = 100, b = 60, t = 20)
  )
```

`r fig5_caption`
:::


# Figure S1. Distribution of α-Gal sIgE results

```{r figS1}
figS1_hist <- figS1_hist %>%
  mutate(Status = factor(Status, levels = c("Negative", "Positive")))

AGAL_TICKS_S1 <- c(0, 0.1, 1, 10, 100)

plot_s1 <- ggplot(figS1_hist, aes(x = bin_centre, y = count, fill = Status)) +
  geom_col(width = 0.2, colour = "black", linewidth = 0.2, alpha = 0.8) +
  geom_vline(
    xintercept = log10(0.1 + 0.1),
    linetype = "dotted",
    colour = "red",
    linewidth = 0.6
  ) +
  scale_x_continuous(
    breaks = log10(AGAL_TICKS_S1 + 0.1),
    labels = AGAL_TICKS_S1
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = comma) +
  scale_fill_manual(values = c(Negative = STEEL, Positive = MiddayBlue)) +
  labs(x = "α-Gal sIgE (kU/L)", y = "Count", fill = "Status") +
  theme_minimal()

ggplotly(plot_s1, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 80, r = 100, b = 60, t = 20)
  )
```

`r figS1_caption`

# Figure S2. Age distribution by sex

::: panel-tabset
## A. All tested

```{r figS2a}
S2_MALE <- BLUEBERRY
S2_FEMALE <- TEAL

plot_s2a <- ggplot(figS2_all, aes(x = Age_cat, y = n, fill = Sex)) +
  geom_col(
    alpha = 0.7,
    position = position_dodge(width = 0.85),
    width = 0.8
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.85),
    width = 0.25,
    linewidth = 0.4,
    colour = "grey30"
  ) +
  scale_fill_manual(
    values = c("M" = S2_MALE, "F" = S2_FEMALE),
    labels = c("M" = "Male", "F" = "Female")
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1800)) +
  labs(x = NULL, y = "Count", fill = "Sex") +
  theme_minimal()

ggplotly(plot_s2a, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 80, r = 100, b = 60, t = 20)
  )
```

`r figS2_caption`

## B. α-Gal sIgE positive

```{r figS2b}
plot_s2b <- ggplot(figS2_pos, aes(x = Age_cat, y = n, fill = Sex)) +
  geom_col(
    alpha = 0.7,
    position = position_dodge(width = 0.85),
    width = 0.8
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.85),
    width = 0.25,
    linewidth = 0.4,
    colour = "grey30"
  ) +
  scale_fill_manual(
    values = c("M" = S2_MALE, "F" = S2_FEMALE),
    labels = c("M" = "Male", "F" = "Female")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Age category", y = "Count", fill = "Sex") +
  theme_minimal()

ggplotly(plot_s2b, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 80, r = 100, b = 60, t = 20)
  )
```

`r figS2_caption`
:::

# Figure S3. Positivity ratios by age and sex

::: panel-tabset
## A. Male vs female

```{r figS3a}
rr_local <- rr_sex_by_age %>%
  mutate(
    Age_cat = factor(Age_cat, levels = age_order),
    significant = ifelse(P_value < 0.05, "Significant", "Not Significant")
  )

plot_s3a <- ggplot(rr_local, aes(x = RR, y = Age_cat)) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "red",
    linewidth = 0.8
  ) +
  geom_point(aes(shape = significant), color = "grey20", size = 3) +
  geom_errorbarh(
    aes(xmin = Lower_CI, xmax = Upper_CI),
    color = "grey20",
    height = 0.3,
    linewidth = 0.6
  ) +
  scale_shape_manual(
    values = c("Significant" = 16, "Not Significant" = 1)
  ) +
  scale_y_discrete(limits = rev) +
  labs(
    x = "Male positivity ratio (vs females in the same age group)",
    y = "Age category"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(plot_s3a, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 60, t = 20))
```

`r figS3_caption`

## B. Age vs 25-34 reference

```{r figS3b}
age_only_order <- c("0-14", "15-24", "35-44", "45-54", "55-64", "65-74", "75-84", "85+")

rr_local <- rr_age_only %>%
  mutate(
    Age_cat = factor(Age_cat, levels = age_only_order),
    significant = ifelse(P_value < 0.05, "Significant", "Not Significant")
  )

plot_s3b <- ggplot(rr_local, aes(x = RR, y = Age_cat)) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "red",
    linewidth = 0.8
  ) +
  geom_point(aes(shape = significant), color = "grey20", size = 3) +
  geom_errorbarh(
    aes(xmin = Lower_CI, xmax = Upper_CI),
    color = "grey20",
    height = 0.3,
    linewidth = 0.6
  ) +
  scale_shape_manual(
    values = c("Significant" = 16, "Not Significant" = 1)
  ) +
  scale_y_discrete(limits = rev) +
  labs(
    x = "Positivity ratio by age group (vs 25-34)",
    y = "Age category"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(plot_s3b, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 60, t = 20))
```

`r figS3_caption`

## C. Age-sex vs female 25-34

```{r figS3c}
S3_MALE <- "#0072B2"
S3_FEMALE <- "#D55E00"

plot_c_data <- combined_rr %>%
  mutate(
    group_label = paste0(Age_cat, " ", Sex),
    significant = ifelse(P_value < 0.05, "Significant", "Not Significant"),
    Age_cat = factor(Age_cat, levels = age_order),
    Sex = factor(Sex, levels = c("M", "F")),
    Sex_label = case_when(
      Sex == "M" ~ "Male",
      Sex == "F" ~ "Female",
      TRUE ~ as.character(Sex)
    )
  ) %>%
  arrange(Age_cat, Sex) %>%
  mutate(group_label = factor(group_label, levels = unique(group_label)))

plot_s3c <- ggplot(plot_c_data, aes(x = RR, y = group_label)) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "red",
    linewidth = 0.8
  ) +
  geom_point(aes(color = Sex_label, shape = significant), size = 3) +
  geom_errorbarh(
    aes(xmin = Lower_CI, xmax = Upper_CI, color = Sex_label),
    height = 0.3,
    linewidth = 0.6
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Male" = S3_MALE, "Female" = S3_FEMALE)
  ) +
  scale_shape_manual(
    values = c("Significant" = 16, "Not Significant" = 1)
  ) +
  scale_y_discrete(limits = rev) +
  labs(
    x = "Positivity ratio of age-sex combinations (vs females 25-34)",
    y = "Age-sex category"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  guides(shape = "none")

ggplotly(plot_s3c, tooltip = c("x", "y", "color")) %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 100, r = 100, b = 60, t = 20)
  )
```

`r figS3_caption`
:::

# Figure S4. Sensitivity of the testing-volume split to the attribution rule

```{r figS4}
rules_levels <- c("Geographic expansion", "Interaction", "Testing intensity")

rules_long <- figS4_rules %>%
  pivot_longer(
    c(geographic_pct, intensity_pct, interaction_pct),
    names_to = "component",
    values_to = "pct"
  ) %>%
  mutate(
    component = recode(
      component,
      geographic_pct = "Geographic expansion",
      intensity_pct = "Testing intensity",
      interaction_pct = "Interaction"
    ),
    component = factor(component, levels = rules_levels),
    rule = factor(rule, levels = rev(figS4_rules$rule))
  )

plot_s4 <- ggplot(rules_long, aes(x = pct, y = rule, fill = component)) +
  geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
  geom_vline(xintercept = c(0, 100), linewidth = 0.3, colour = "grey30") +
  scale_fill_manual(
    values = c(
      "Geographic expansion" = "grey60",
      "Interaction" = "grey85",
      "Testing intensity" = "grey20"
    )
  ) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Percentage of observed 2014-2024 growth in testing volume",
    y = NULL,
    fill = NULL
  ) +
  theme_minimal()

ggplotly(plot_s4, tooltip = c("x", "y", "fill")) %>%
  layout(
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.2),
    margin = list(l = 300, r = 40, b = 90, t = 20)
  )
```

`r figS4_caption`

# Figure S5. Patient-to-clinician distance

```{r figS5}
median_dist <- median(figS5_distances$distance_km, na.rm = TRUE)

plot_s5 <- ggplot(figS5_distances, aes(x = distance_km)) +
  # boundary = 0 so the first bin is 0-6 km rather than straddling zero: 1,778
  # of these pairs share a postcode and would otherwise fall half outside the
  # panel. The x range is a zoom (coord_cartesian), not a filter, so no bar is
  # silently dropped.
  geom_histogram(
    binwidth = 6,
    boundary = 0,
    fill = MidnightBlue,
    alpha = 0.7,
    colour = "white"
  ) +
  geom_vline(
    xintercept = median_dist,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.5
  ) +
  coord_cartesian(xlim = c(0, 300)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = paste0("Distance (km); median ", round(median_dist, 1), " km"),
    y = "Number of individuals"
  ) +
  theme_minimal()

ggplotly(plot_s5, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 80, r = 40, b = 60, t = 20))
```

`r figS5_caption`

# Figure S6. Geographic concentration standardised by testing volume

::: panel-tabset
## A. Lorenz curves vs population

```{r figS6a}
supp_theme <- theme_minimal() +
  theme(
    axis.text = element_text(size = 9),
    axis.title = element_text(size = 10)
  )

L_CASES <- "α-Gal sensitised individuals (denominator: population)"
L_TESTED <- "Persons tested (denominator: population)"

case_col <- "#18444E"
test_col <- ORANGE

lorenz_pop <- figS6_lorenz %>%
  filter(measure != "cases | tested") %>%
  mutate(measure = factor(measure, levels = c(L_CASES, L_TESTED)))

plot_s6a <- ggplot(
  lorenz_pop,
  aes(x = cum_denom * 100, y = cum_event * 100, colour = measure)
) +
  geom_abline(
    intercept = 0, slope = 1,
    linetype = "dashed", colour = "red", linewidth = 0.4
  ) +
  geom_line(linewidth = 0.8) +
  scale_colour_manual(values = c(case_col, test_col), name = NULL) +
  scale_x_continuous(
    breaks = seq(0, 100, 20), limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    breaks = seq(0, 100, 25), limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  annotate(
    "text", x = 97, y = 26, hjust = 1, size = 3.2,
    label = paste0(
      "Gini, sensitised = ", round(figS6_gini$g_cases_pop, 3),
      "\nGini, persons tested = ", round(figS6_gini$g_tested_pop, 3)
    )
  ) +
  labs(
    x = "Cumulative percentage of population (regions ranked by rate, highest first)",
    y = "Cumulative percentage of events"
  ) +
  supp_theme

ggplotly(plot_s6a, tooltip = c("x", "y", "colour")) %>%
  layout(
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.2),
    margin = list(l = 90, r = 40, b = 110, t = 20)
  )
```

`r figS6_caption`

## B. Lorenz curve vs testing volume

```{r figS6b}
plot_s6b <- ggplot(
  figS6_lorenz %>% filter(measure == "cases | tested"),
  aes(x = cum_denom * 100, y = cum_event * 100)
) +
  geom_abline(
    intercept = 0, slope = 1,
    linetype = "dashed", colour = "red", linewidth = 0.4
  ) +
  geom_line(colour = case_col, linewidth = 0.8) +
  scale_x_continuous(
    breaks = seq(0, 100, 20), limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    breaks = seq(0, 100, 25), limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  annotate(
    "text", x = 97, y = 8, hjust = 1, size = 3.2,
    label = paste0(
      "Concentration index of α-Gal sensitised\nagainst testing volume = ",
      round(figS6_gini$g_cases_tested, 3)
    )
  ) +
  labs(
    x = "Cumulative percentage of persons tested (regions ranked by positivity, highest first)",
    y = "Cumulative percentage of sensitised individuals"
  ) +
  supp_theme

ggplotly(plot_s6b, tooltip = c("x", "y")) %>%
  layout(showlegend = FALSE, margin = list(l = 90, r = 40, b = 70, t = 20))
```

`r figS6_caption`

## C. Standardised positivity ratios

```{r figS6c}
#| fig-height: 8

spr_top <- figS6_supp %>%
  arrange(desc(case_rate_1M)) %>%
  slice_head(n = 30) %>%
  mutate(SA_name = fct_reorder(SA_name, spr))

plot_s6c <- ggplot(spr_top, aes(x = spr, y = SA_name)) +
  geom_vline(
    xintercept = 1, linetype = "dashed", colour = "red", linewidth = 0.4
  ) +
  geom_errorbar(
    aes(xmin = spr_lo, xmax = spr_hi),
    orientation = "y",
    width = 0,
    linewidth = 0.4,
    colour = "grey40"
  ) +
  geom_point(
    aes(
      size = tested,
      text = paste0(
        SA_name, " (", State, ")",
        "<br>SPR ", round(spr, 2),
        " (95% CI ", round(spr_lo, 2), "-", round(spr_hi, 2), ")",
        "<br>", cases, " sensitised of ", tested, " tested"
      )
    ),
    colour = case_col,
    fill = case_col,
    shape = 21
  ) +
  scale_size_continuous(range = c(1, 4), name = "Persons\ntested") +
  scale_x_continuous(transform = "log2", breaks = c(0.25, 0.5, 1, 2, 4)) +
  labs(
    x = "Standardised positivity ratio (observed / expected given testing volume)",
    y = NULL
  ) +
  supp_theme +
  theme(axis.text.y = element_text(size = 8))

ggplotly(plot_s6c, tooltip = "text") %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 170, r = 110, b = 70, t = 20)
  )
```

`r figS6_caption`

## D. Sensitisation vs testing rate

```{r figS6d}
#| fig-height: 7

# The published panel labels the 12 highest-burden regions with repelled text,
# which ggplotly cannot reproduce; here every point names itself on hover.
scatter_data <- figS6_supp %>%
  filter(tested > 0, case_rate_1M > 0) %>%
  mutate(
    Region = ifelse(is_top30, "Top 30 by rate", "Other regions"),
    Region = factor(Region, levels = c("Other regions", "Top 30 by rate"))
  )

iso_line <- tibble(
  test_rate_1M = range(scatter_data$test_rate_1M, na.rm = TRUE)
) %>%
  mutate(case_rate_1M = figS6_gini$nat_positivity * test_rate_1M)

plot_s6d <- ggplot(scatter_data, aes(x = test_rate_1M, y = case_rate_1M)) +
  geom_line(
    data = iso_line, colour = "red", linetype = "dashed", linewidth = 0.4
  ) +
  geom_point(
    aes(
      colour = Region,
      size = tested,
      text = paste0(
        SA_name, " (", State, ")",
        "<br>", round(case_rate_1M, 1), " sensitised per 1M PPY",
        "<br>", round(test_rate_1M, 1), " tested per 1M PPY",
        "<br>positivity ", round(positivity * 100, 1), "%"
      )
    ),
    alpha = 0.75,
    shape = 16
  ) +
  scale_colour_manual(
    values = c("Other regions" = MidnightBlue, "Top 30 by rate" = MiddayBlue),
    name = NULL
  ) +
  scale_size_continuous(range = c(1, 5), name = "Persons\ntested") +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  labs(
    x = "Persons tested per 1M PPY (log scale)",
    y = "α-Gal sensitised individuals per 1M PPY (log scale)"
  ) +
  supp_theme

ggplotly(plot_s6d, tooltip = "text") %>%
  layout(
    legend = list(
      orientation = "v",
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    margin = list(l = 90, r = 120, b = 70, t = 20)
  )
```

`r figS6_caption`
:::

# Figure S7. Population- and testing-standardised burden

Each panel is the published static pair: all of mainland Australia on the left,
south-east Queensland and eastern New South Wales on the right. Figure 3 above
is the zoomable equivalent of panel A.

```{r figS7-setup, include=FALSE}
map_sf <- sa3_sf %>%
  select(SA_code, SA_name, geometry) %>%
  mutate(SA_code = as.numeric(as.character(SA_code))) %>%
  left_join(
    figS6_supp %>%
      select(SA_code, case_rate_1M, test_rate_1M, positivity_mapped),
    by = "SA_code"
  )

map_xlim <- c(112, 154.5)
map_ylim <- c(-44, -9.5)
map_xlim_ec <- c(147, 154.5)
map_ylim_ec <- c(-39, -25)

base_map_theme <- theme_bw() +
  theme(
    plot.title = element_text(size = 10, face = "bold"),
    plot.subtitle = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7.5),
    legend.key.width = unit(3.5, "mm"),
    legend.key.height = unit(8, "mm"),
    legend.position = "right",
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

choropleth <- function(fill_var, title, subtitle, legend_title,
                       trans = "identity", east_coast = FALSE) {
  ggplot(map_sf) +
    geom_sf(
      aes(fill = .data[[fill_var]]),
      colour = "grey70",
      linewidth = 0.03
    ) +
    scale_fill_viridis_c(
      option = "magma",
      direction = -1,
      transform = trans,
      na.value = "grey93",
      name = legend_title,
      labels = comma
    ) +
    coord_sf(
      xlim = if (east_coast) map_xlim_ec else map_xlim,
      ylim = if (east_coast) map_ylim_ec else map_ylim,
      expand = FALSE
    ) +
    labs(title = title, subtitle = subtitle) +
    base_map_theme +
    theme(
      # The inset shares its national panel's fill scale, so a second legend
      # would only duplicate it
      legend.position = if (east_coast) "right" else "none",
      plot.margin = if (east_coast) {
        margin(2, 0, 2, 0, "mm")
      } else {
        margin(2, 1, 2, 0, "mm")
      }
    )
}

nat_pos_pct <- round(figS6_gini$nat_positivity * 100, 1)
```

::: panel-tabset
## A. Sensitisation rate

```{r figS7a}
#| fig-width: 10
#| fig-height: 6

(
  choropleth(
    "case_rate_1M",
    "A. Population-standardised sensitisation rate",
    "α-Gal sensitised individuals per 1M person-years, 2014-2024",
    "Sensitised\nper 1M PPY",
    trans = "sqrt"
  ) +
    choropleth(
      "case_rate_1M", "", "", "Sensitised\nper 1M PPY",
      trans = "sqrt", east_coast = TRUE
    )
) +
  plot_layout(widths = c(1, 0.30))
```

`r figS7_caption`

## B. Testing intensity

```{r figS7b}
#| fig-width: 10
#| fig-height: 6

(
  choropleth(
    "test_rate_1M",
    "B. Testing intensity",
    "Persons tested per 1M person-years, 2014-2024",
    "Tested per\n1M PPY",
    trans = "sqrt"
  ) +
    choropleth(
      "test_rate_1M", "", "", "Tested per\n1M PPY",
      trans = "sqrt", east_coast = TRUE
    )
) +
  plot_layout(widths = c(1, 0.30))
```

`r figS7_caption`

## C. Testing-standardised burden

```{r figS7c}
#| fig-width: 10
#| fig-height: 6

(
  choropleth(
    "positivity_mapped",
    "C. Testing-standardised burden",
    paste0(
      "Test positivity (empirical Bayes smoothed), regions with >= ",
      figS6_gini$min_tested,
      " persons tested;\nnational value ", nat_pos_pct,
      "%. Grey = insufficient testing"
    ),
    "Positivity"
  ) +
    choropleth("positivity_mapped", "", "", "Positivity", east_coast = TRUE)
) +
  plot_layout(widths = c(1, 0.30))
```

`r figS7_caption`
:::
